---
title: "Lab Decision Tree"
author: "Tiffany Wibisono"
date: "2022-10-26"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: united
    highlight: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rpart)
library(rpart.plot)
library(forecast)
library(caret)
```

# Regression Tree
## 1. Load Data
```{r}
toyota <- read.csv("ToyotaCorolla.csv", header = TRUE)

head(toyota)
str(toyota)
names(toyota)
nrow(toyota)

#select necessary variables only
toyota = toyota[, c(3:4, 7:9, 12, 14, 17, 19, 21, 25:26, 28, 30, 34, 39)]
names(toyota)
t(t(names(toyota)))
str(toyota)

#set the target variable as a factor
toyota$Fuel_Type = as.factor(toyota$Fuel_Type)
```
## 2. Set Training and Validation Set
```{r}
set.seed(666)

train_index = sample(1:nrow(toyota), 0.6*nrow(toyota))
valid_index = setdiff(1:nrow(toyota), train_index)

train_df = toyota[train_index, ]
valid_df = toyota[valid_index, ]

#checking
head(train_df)
head(valid_df)
str(train_df)
str(valid_df)
nrow(train_df)
nrow(valid_df)
```
## 3. Build Decision Tree
```{r}
regress_tr = rpart( Price ~.,
                   data = train_df, method = "anova", maxdepth = 3)
prp(regress_tr)

#Predict training set
predict_train = predict(regress_tr, train_df)
accuracy(predict_train, train_df$Price)

#Predict validation set
predict_valid = predict(regress_tr, valid_df)
accuracy(predict_valid, valid_df$Price)
```
## 4. Predict New Record
```{r}
#Create new dataframe
new_record <- data.frame(Age_08_04 = 77, 
                         KM = 117000, 
                         Fuel_Type = "Petrol", 
                         HP = 110, 
                         Automatic = 0, 
                         Doors = 5, 
                         Quarterly_Tax = 100, 
                         Mfr_Guarantee = 0, 
                         Guarantee_Period = 3, 
                         Airco = 1, 
                         Automatic_airco = 0, 
                         CD_Player = 0, 
                         Powered_Windows = 0, 
                         Sport_Model = 0, 
                         Tow_Bar = 1)


#Predict new record
regress_tr_pred = predict(regress_tr, newdata = new_record)
regress_tr_pred
```
## 5. Summary
- RMSE in the training set is 1295.696, while RMSE in the validation set is 1581.903. It is normal to have a slightly higher RMSE in the validation set than in the training set; thus, in this case, the model looks good.
- The predicted price of the new record would be $7969.59.


# Classification Tree
## 1. Binning
```{r}
#Converting the Price to a categorical variable
toyota$cat_price <- ifelse(toyota$Price <= mean(toyota$Price, na.rm = TRUE), "0", "1")
table(toyota$cat_price)

toyota$cat_price <- as.factor(toyota$cat_price)


#Remove the numerical Price variable to avoid confusion 
toyota_cat <- toyota[,- c(1)]
names(toyota_cat)
str(toyota_cat)
```
## 2. Set Training and Validation Set
```{r}
set.seed(666)

train_cat_index = sample(1:nrow(toyota_cat), 0.6*nrow(toyota_cat))
valid_cat_index = setdiff(1:nrow(toyota_cat), train_cat_index)

train_cat_df = toyota_cat[train_cat_index, ]
valid_cat_df = toyota_cat[valid_cat_index, ]

head(train_cat_df)
head(valid_cat_df)
str(train_cat_df)
str(valid_cat_df)
nrow(train_cat_df)
nrow(valid_cat_df)
```
## 3. Build Classification Tree
```{r}
names(train_cat_df)

class_tr = rpart(cat_price ~.,
                 data = train_cat_df, method = "class", maxdepth = 3)
prp(class_tr, cex = 1, tweak = 1)
```
## 4. Model Evaluation
### 4.1 Confusion Matrix
```{r}
#Training set
class_tr_train_predict = predict(class_tr, train_cat_df,
                                 type = "class")
t(t(head(class_tr_train_predict, 10)))

confusionMatrix(class_tr_train_predict, train_cat_df$cat_price)

#Validation set
class_tr_valid_predict = predict(class_tr, valid_cat_df,
                                 type = "class")
t(t(head(class_tr_valid_predict, 10)))
confusionMatrix(class_tr_valid_predict, valid_cat_df$cat_price)

#Compute the probabilities
class_tr_valid_predict_prob = predict(class_tr, valid_df,
                                      type = "prob")
head(class_tr_valid_predict_prob)
```
### 4.2 ROC Curve
```{r}
library(ROSE)

ROSE:: roc.curve(valid_cat_df$cat_price, class_tr_valid_predict)
```
## 5. Predict New Record
```{r}
cat_price_class_tr = predict(class_tr, newdata = new_record)
cat_price_class_tr
```
## 6. Summary
- The top predictor in the classification tree is the Age_08_0>=57.
- The accuracy in the training set is 0.9001, and the accuracy in the validation set is 0.8904. It is normal to have a slightly lower accuracy in the validation than in the training set; thus, in this case, the model is good.

# Compare the Trees
## 1. Calculate the Average Price 
```{r}
mean(toyota$Price)
```
## 2. Summary
- The average price of used Toyota cars is $10,730.82.
- Using a regression tree, the predicted price of a new data is $7969.59, which is 25.73% lower than the average price. 
-Using a classification tree, the probability of the predicted price of a new data to be lower than the average price is 0.95. The result tells us that the price of the new data is most likely to be lower than the average price. There is only a 4% chance that it is going to be priced above the average price.
- Based on the result of regression and classification tree, it shows that the new data is likely to be priced below $10,730.82.
