---
title: "Lab Clustering"
author: "Tiffany Wibisono"
date: "2022-11-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("factoextra")
library(factoextra)
```
## 1. Data Cleaning
```{r}
univ <- read.csv("Universities.csv", header = TRUE)

#remove records with missing data
univ_complete <- na.omit(univ)

```
##2. Hierarchical clustering using normalised data, complete linkage and Manhattan
### 2.1 Remove categorical variables and ID
```{r}
#These are college name, state and public vs private
univ_cont <- univ_complete[, -c(1:3)]

t(t(names(univ_cont)))
```
### 3.2 Normalize Data
```{r}
univ_cont_norm <- sapply(univ_cont, scale)
```
### 3.3 Compute the normalized distance matrix using Manhattan distance
```{r}
distance_matrix <- dist(univ_cont_norm, method = "manhattan")
distance_matrix
```
### 3.4 Create hierarchical clusters
```{r}
#syntax
hierarchical_cluster <- hclust(distance_matrix, method = "complete")
hierarchical_cluster

#plot the dendrogram
plot(hierarchical_cluster, hang = "-200", ann = TRUE)

#check cluster quality
library(factoextra)

fviz_nbclust(univ_cont_norm, 
             hcut, method = "silhouette") +
  labs(subtitle = "Silhouette method")
```
### 3.5 Compute cluster membership using k = 5
```{r}
hierarchical_cluster_k5 = cutree(hierarchical_cluster, k = 5)

hierarchical_cluster_k5
```
### 3.6 Create a data frame of the cluster membership
```{r}
hierarchical_cluster_k5_df = as.data.frame(hierarchical_cluster_k5)

head(hierarchical_cluster_k5_df)

#rename the the column title
names(hierarchical_cluster_k5_df)[1] = "Cluster"
tail(hierarchical_cluster_k5_df)

#count the number of frequency in each cluster
table(hierarchical_cluster_k5_df$Cluster)
```
### 3.7 Merge with Original Data
```{r}
univ_cont_w_cluster = cbind(univ_cont, hierarchical_cluster_k5_df)

head(univ_cont_w_cluster)

names(univ_cont_w_cluster)
```
## 4. Investigate Cluster Characteristics
```{r}
#cluster characteristics, numerical variables

ncol(univ_cont_w_cluster)
t(t(names(univ_cont)))

univ_cont_w_cluster_df = aggregate(univ_cont[, c(1:17)],
                                   by = univ_cont_w_cluster[18],
                                   FUN = mean)

#cluster characteristics, categorical variables
table(univ_cont_w_cluster_df$Cluster)
```