---
title: Lab kNN
author: "Tiffany Wibisono"
date: "2022-10-14"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: united
    highlight: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(caret)
```


## 1. Load Data
```{r}
bank <- read.csv("UniversalBank.csv")
head(bank, 10)
str(bank)
names(bank)
```
### 1.1 Data Clean Up
```{r}
#remove ID and zip code column
bank = bank[,-c(1,5)]
names(bank)

#Reorder variables. Put the response last
bank = bank[, c(1:7, 9:12, 8)]
head(bank)
str(bank)

#Set categorical variable
bank$Education = as.factor(bank$Education)
bank$Securities.Account = as.factor(bank$Securities.Account)
bank$CD.Account = as.factor(bank$CD.Account)
bank$Online = as.factor(bank$Online)
bank$CreditCard = as.factor(bank$CreditCard)

#Rename outcome variable values
bank$Personal.Loan = factor(bank$Personal.Loan,
                            levels = c("0", "1"),
                            labels = c("No", "Yes"))
table(bank$Personal.Loan)
```
## 2. Create Training and Validation Set
```{r}
set.seed(666)

train_index = sample(1:nrow(bank), 0.6*nrow(bank))
valid_index = setdiff(1:nrow(bank), train_index)

train_df = bank[train_index,]
valid_df = bank[valid_index,]

nrow(train_df)
nrow(valid_df)
```
## 3. Define New Customer
```{r}
#Create new data set for new customer
new_cust = data.frame(Age = 40,
                      Experience = 10,
                      Income = 84,
                      Family = 2,
                      CCAvg = 2,
                      Education = 2,
                      Mortgage = 0,
                      Securities.Account = 0,
                      CD.Account = 0,
                      Online = 1,
                      CreditCard = 1)
new_cust

#Set categorical variable as factor
new_cust$Education = as.factor(new_cust$Education)
new_cust$Securities.Account = as.factor(new_cust$Securities.Account)
new_cust$CD.Account = as.factor(new_cust$CD.Account)
new_cust$Online = as.factor(new_cust$Online)
new_cust$CreditCard = as.factor(new_cust$CreditCard)

new_cust
```
## 4. Prepare for kNN
```{r}
#Normalization, only for numerical variable
train_norm = train_df
valid_norm = valid_df

#Create a normalizing algorithm using all variables except the categorical variables
norm_values = preProcess(train_df[-c(12)],
                         method = c("center",
                                    "scale"))
#Predict the normalized value of the training set
train_norm[, -c(12)] = predict(norm_values,
                               train_df[, -c(12)])
head(train_norm)

#Predict the normalized value of the validation set
valid_norm[, -c(12)] = predict(norm_values,
                               valid_df[, -c(12)])
head(valid_norm)

#Predict the normalized value for the new customers 
new_cust_norm = predict(norm_values, new_cust)
new_cust_norm
```
## 5. Train the kNN for Predictions
### 5.1 Train and Predict when k = 3
```{r}
knn_model_k3 <- caret::knn3(Personal.Loan ~ ., 
                            data = train_norm, k = 3)
knn_model_k3

#Predict the training set
knn_pred_k3_train <- predict(knn_model_k3, 
                             newdata = train_norm[, -c(12)], 
                             type = "class")
head(knn_pred_k3_train)

#Evaluate 
confusionMatrix(knn_pred_k3_train, as.factor(train_norm[, 12]),
                positive = "Yes")
```
### 5.2 Train and Predict when k = 5
```{r}
knn_model_k5 <- caret::knn3(Personal.Loan ~ ., 
                            data = train_norm, k = 5)
knn_model_k5

#Predict the training set
knn_pred_k5_train = predict(knn_model_k5,
                            newdata = train_norm[, -c(12)],
                            type = "class")
head(knn_pred_k5_train)

#Evaluate
confusionMatrix(knn_pred_k5_train,as.factor(train_norm[, 12]),
                positive = "Yes")
```
### 5.3 Train and Predict when k =7
```{r}
knn_model_k7 <- caret::knn3(Personal.Loan ~ ., 
                            data = train_norm, k = 7)
knn_model_k7

#Predict the training set
knn_pred_k7_train = predict(knn_model_k7,
                            newdata = train_norm[, -c(12)],
                            type = "class")
head(knn_pred_k5_train)

#Evaluate
confusionMatrix(knn_pred_k7_train,as.factor(train_norm[, 12]),
                positive = "Yes")
```
## 6. Predict Validation Set
### 6.1 Predict Validation Set when k = 3
```{r}
#Predict the validation set
knn_pred_k3_valid = predict(knn_model_k3, 
                            newdata = valid_norm[, -c(12)],
                            type = "class")
head(knn_pred_k3_valid)

# Evaluate
confusionMatrix(knn_pred_k3_valid, as.factor(valid_norm[, 12]),
                positive = "Yes")
```
### 6.2 Predict Validation Set when k = 5
```{r}
#Predict the validation set
knn_pred_k5_valid = predict(knn_model_k5, 
                            newdata = valid_norm[, -c(12)],
                            type = "class")
head(knn_pred_k5_valid)

#Evaluate
confusionMatrix(knn_pred_k5_valid, as.factor(valid_norm[, 12]),
                                            positive = "Yes")
```
### 6.3 Predict validation Set when k = 7
```{r}
#Predict the validation set
knn_pred_k7_valid = predict(knn_model_k7,
                            newdata = valid_norm[, -c(12)],
                            type = "class")
head(knn_pred_k7_valid)

#Evaluate
confusionMatrix(knn_pred_k7_valid, as.factor(valid_norm[, 12]),
                positive = "Yes")
```
## 7. Decision for Personal Loan using New Data
```{r}
new_cust_predict = predict(knn_model_k3, newdata = new_cust_norm,
                           type = "class")
new_cust_predict
```
## 8. Model Evaluation
```{r}
library(ROSE)
ROSE::roc.curve(valid_norm$Personal.Loan, knn_pred_k3_valid)
```
## 9. Summary
- Based on the prediction, it shows that model 3 gives us the best result. The accuracy in the training set is 97.53%, and the accuracy in the validation set is 96.15%.
- When I classify a new customer, it is predicted that she will not accept the loan offer.